Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure

### 2017-04-05 documents/TriangleScenarioArXiv/ts.tex
typos in abstract/authors multiple lines

% === Revtex Declaration ===
\documentclass[aps, 10pt, english, twoside, pra, nofootinbib, tightenlines, longbibliography]{revtex4-1}

@@

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
%    \email{ewolfe@perimeterinstitute.ca}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed nonlocal, although they are more accurately termed non-classical, as the presence of correlations certifies the genuinely non-classical nature of a causal setup in a device-independent fashion.  Whenever a quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as non-classical. Any constraint which is satisfied by all correlations which \textit{are} classically compatible with a given causal scenario defines a causal compatibility criterion. Such criteria were recently were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms -- such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} -- scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols --- a popular example being Quantum Key Distribution~\cite{Bennett_2014} -- are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. History, every opportunity of quantum phenomena which fails to be emulated by classical resources has led to a practical exploitation of the nonclassical feature in order to solve a some computational of communicational problem~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is the discovery of new situations wherein quantum mechanics offers an advantage, and certification that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenemena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{nonclassicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatability inequalities, such as Bell of Instrumental inequalities, charechterize the space of observations that are compatible with a hypothesized causal structure, albeit the charechterization offered by practically deriveable inequalities often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_scenario}. The fundamental Bell structure involves noncommunicating parties making measurements on some hidden shared resource $\la$, where the measuremement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum nonclassicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex scenarios, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal scenario: the Triangle structure (\cref{fig:triangle_structure}).

@@

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum nonclassicality. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of nonclassicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first concieved of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the Triangle structure that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:found_inequalities} (specifically \cref{eq:ww_ineq}) as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

>>>>

% === Revtex Declaration ===
\documentclass[aps, 10pt, english, twoside, pra, nofootinbib, tightenlines, longbibliography, superscriptaddress]{revtex4-1}

@@

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    \affiliation{University of Waterloo, Waterloo, Ontario, Canada, N2L 3G1}
    \author{Elie Wolfe}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada, N2L 2Y5}
    % \email{ewolfe@perimeterinstitute.ca}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed nonlocal, although they are more accurately termed non-classical, as the presence of correlations certifies the genuinely non-classical nature of a causal setup in a device-independent fashion.  Whenever a quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as non-classical. Any constraint which is satisfied by all correlations which \textit{are} classically compatible with a given causal scenario defines a causal compatibility criterion. Such criteria were recently were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution in this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \textit{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear. In the realm of computation, quantum algorithms -- such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} -- scale exponentially better than their classical counterparts. In the realm of secure communication, quantum protocols --- a popular example being Quantum Key Distribution~\cite{Bennett_2014} -- are able to provide privacy even against hypothetical adversaries with unlimited computational power, a desideratum which classical protocols are unable to fulfill. History, every opportunity of quantum phenomena which fails to be emulated by classical resources has led to a practical exploitation of the non-classical feature in order to solve a some computational of communicational problem~\cite{Neilsen_Chaung_2011}. Motivated by past successes, a primal objective of modern quantum information theory is the discovery of new situations wherein quantum mechanics offers an advantage, and certification that the quantum advantage is genuine.

    From a foundational prospective, the most robust demonstrations of quantum phenomena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{non-classicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatibility inequalities, such as Bell of Instrumental inequalities, characterizer the space of observations that are compatible with a hypothesized causal structure, albeit the characterization offered by practically derivable inequalities often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_scenario}. The fundamental Bell structure involves non-communicating parties making measurements on some hidden shared resource $\la$, where the measurement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum non-classicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex scenarios, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal scenario: the Triangle structure (\cref{fig:triangle_structure}).

@@

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum non-classicality. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of non-classicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first conceived of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the Triangle structure that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:found_inequalities} (specifically \cref{eq:ww_ineq}) as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.
